Hyperplane Elimination for Quickly Enumerating Local Optima

### 2015-09-15 Hyperplane Enumeration/main.tex
Filled out more r-bit local optima stuff

\subsection{Examining Local Optima}

\begin{figure*}
  \centering
  \includegraphicsfit{length-radius}
  \caption{Number of $r$-bit local optima as $N$ increases for each problem.}
  \label{fig-length-radius}
\end{figure*}

TODO Consider discussing neutral networks

\section{Conclusions}

>>>>

Increasing $r$ also results is a substantial reduction in the
total number of local optima, as shown in Figure~\ref{fig-length-radius}.
For every problem except Concatenated Traps, increasing $r$ not only reduces the
number of local optima, it decreases the growth rate with respect to $N$. As there
are fewer local optima to enumerate, Algorithm~\ref{alg-enumerate} requires less
time to enumerate them all. In Concatenated Traps each trap requires exactly $k$ flips in
order to move between local optima. Therefore, while $r < k$ the number of local optima
cannot change and once $r \ge k$ there is exactly 1 local optima.

Previous work~\cite{ochoa:2015:crossovernetworks} noted that counter-intuitively Adjacent
NKq has more local optima than Random NKq for $r=1$ and $N=20$, even though the former is generally
easier to solve. However, many of these optima are ``weak'' in the sense that they are not also
2-bit local optima. Here we extend those conclusions for up to $N=59$ and up to $r=4$. Adjacent
NKq tend to have more 1-bit local optima than Random NKq, most of which disappear as $r$
increases. Furthermore, the growth rate of local optima between the two landscapes diverges
as $r$ increases.

MAX-kSAT and Ising Spin Glasses generally appear to contain far more local optima than
any of the other problems we enumerated. With just $N=49$, Ising Spin Glasses average
nearly 800 million local optima. For comparison, Adjacent NKq with $N=49$ averages fewer than 300 thousand,
and with $N=78$ Adjacent NKq still only has fewer than 500 million.
As with NKq, the polynomially solvable Ising Spin
Glass problem had more 1-bit local optima than the NP-Hard MAX-kSAT. Unlike with NKq, this
relationship generally does not change as $r$ increases. Even for $r=4$ the general
number and growth rate of local optima appears to stay the same for both problems.

\begin{figure*}
  \centering
  \includegraphicsfit{length-radius}
  \caption{Number of $r$-bit local optima for each size of each problem. Note that
           the number of local optima in Concatenated Traps does not change while $r<k$.}
  \label{fig-length-radius}
\end{figure*}

In our original experiment we tested problems up to size $N=100$, which was sufficient to
cause all three methods using $r=1$ to fail all problems except for Hyper-Reorder on Concatenated
Traps. However, using higher $r$ values, Hyper-Reorder is able to significantly extend its range.
On Random NKq, Hyper-Reorder was successful at finding 2-bit local optima for $N=73$. On Adjacent
NKq using all $r$ tested except 1 were successful on $N=100$. To push Algorithm~\ref{alg-enumerate}
to the extreme we tested Adjacent NKq with $N=200$ and $r=4$, with all 30 runs successful. On
average these runs required 5.6 minutes to complete. However, there was significant variance
in both the runtime and number of local optima. 18 of 30 runs completed in under 1 minute,
while the slowest run took 1.5 hours. 7 had less than one thousand 4-bit local optima, while 5
had over one hundred thousand.

\section{Conclusions}
TODO Consider discussing neutral networks
